\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{mathrsfs}

%SetFonts

%SetFonts


\title{CSE537 HW2}
\author{Tim Zhang}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

\section{Performance}
\begin{center}
 \begin{tabular}{||c c c c c||} 
 \hline
 Algorithm & Input & Time (ms) & Searches & Arc Prunings \\ [0.5ex] 
 \hline\hline
 DFSB++ & backtrack\_1 & 205.00199999999998 & 146 & 345 \\
 DFSB++ & backtrack\_2 & 754.674 & 368 & 546 \\ 
 DFSB++ & backtrack\_easy & 0.9679999999999964 & 8 & 10 \\
 DFSB++ & backtrack\_hard & 15331.961000000001 & 2679 & 1389 \\ 
 DFSB & backtrack\_1 & Failure & 163957 & 0 \\
 DFSB & backtrack\_2 & Failure & 103541 & 0 \\ 
 DFSB & backtrack\_easy & 0.17800000000000082 & 8 & 0 \\
 DFSB & backtrack\_hard & Failure & 44000 & 0 \\ 
 Min Conflicts & minconflict\_1 & 0.4680000000000035 & 7 & 0 \\
 Min Conflicts & minconflict\_2 & 2445.835 & 100036 & 0 \\ 
 Min Conflicts & minconflict\_easy & 1.3279999999999959 & 13 & 0 \\
 Min Conflicts & minconflict\_hard & Failure & 400000 & 0 \\ [1ex] 
 \hline
\end{tabular}
\end{center}

\section{Observations}
\subsection{DFSB}
Naive DFSB could not solve any of the input sequences under the time constraints whereas under DFSB++ all instances were solvable.  This is due to the massive reduction in the search space given by the heuristics used in DFSB++.  

Effectively each of the heuristics in turn pruned the search tree at branches which were either destined to fail or incompatible to begin with.  In this way both redundant and fruitless branches were avoided completely.

It is notable that for the easy instance of the test input DFSB had superior performance.  This is due to the fact that DFSB++ uses computation time to compute the heuristics even on the simple input case.  However in the more complex cases this extra computation time is more than made up for by the reduction in search space.

\subsection{Min Conflicts}
The observed solution time of Min Conflicts is heavily dependent on the starting configuration of the CSP, which is generated randomly in my program.  For this reason there is heavy variation in observed running times on even the same input instance.

To avoid the normal pitfalls of local search I implemented my algorithm using random restart which after an arbitrary amount of searches the algorithm would terminate and be reinitialized with a new starting configuration and run again.

Due to random restart, the original hard instance "minconflict\_hard" would be solvable in some cases and trivially complete depending on the starting configuration.

\end{document}  